---
title: "Diwali Sales Trend Analysis"
subtitle: "INFO 526 - Project 1"
author: "Group: Diwali Dhamaka"
format: html
editor: visual
execute:
  warning: false
---

```{r}
#| label: load-packages
#| include: false

# Load packages here
pacman::p_load(dplyr,
               broom,
               ggplot2,
               scales,
               kableExtra)

# Plot theme
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 11))
```

## Abstract

Diwali, the Indian festival of lights, holds cultural and economic significance comparable to the Christmas shopping period. Leveraging the diwali_sales_data.csv dataset from the 2023 TidyTuesday repository, this study explores transaction trends during Diwali and spending patterns. Several visualizations were made utilizing the dataset's features to answer two questions:

1)   What is the most prevalent customer group for each product category?

2)   How does age affect a customer's total spending during the festival?

Question 1 was assessed using barplots. Results revealed notable differences, such as a higher number of female customers and increased transactions among singles. Clothing, entertainment, and food dominate transactions, aligning with the festival's festive nature.

To address potential confounding factors for question 1 analysis, a boxplot was generated to examine product category prices. Price ranges did not significantly differ suggesting that prices did not impact transaction numbers.

Question 2 was assessed using both boxplots and regressions. The boxplot shows consistent spending across age groups, with female customers exhibiting more frequent and larger outlier values. Regression analysis reveals that male spending remains consistent across all ages, while female spending peaks around age 50. Further analysis on men's spending indicates age has no significant effect, supported by p-values exceeding $\alpha$ = 0.05.

## Introduction
Diwali is the Indian festival of lights. In addition to its cultural
significance, it is a period of economic activity comparable to the Christmas
shopping period. Due to this economic activity, many businesses would find it
worthwhile to look into sales data to find transaction trends and develop
policies to increase revenue in future festivals.

The diwali_sales_data.csv is one such dataset sourced from the 2023 TidyTuesday
data repository. The .csv spans multiple years and contains 13 variables and
11251 transactions. A customer can have multiple transactions and across
multiple festivals. Of the transactions, 12 transactions did not have
information regarding the transaction amount and are excluded from the data.
Information for customers under the age of 18 were also excluded due to a
significant number of nonsensical entries.

```{r}
#| label: dataset-table
#| echo: false
#| message: false
# Load Diwali dataset
diwali_sales <- read.csv("data/diwali_sales_data.csv")

# Variable type
diwali_class <- c("Double", "Character", "Character", "Character", "Character",
                  "Double", "Double", "Character", "Character", "Character",
                  "Character", "Double", "Double")

# Variable Description
diwali_desc <- c("User ID number",
                 "Customer name",
                 "Product ID number",
                 "Gender (e.g. Male, Female)",
                 "Age group",
                 "Age",
                 "Marital status (e.g. Married, Single)",
                 "State",
                 "Geographic zone",
                 "Occupation",
                 "Category of the product",
                 "Number of orders",
                 "Indian rupees spent by the customer")

# Generate summary values
diwali_values <- c("Seven digit unique number",
                   "Name",
                   "Product Code",
                   "F/M",
                   "0-17, 18-25, 26-35, 36-45, 46-50, 51-55, 55+",
                   "12-92",
                   "0/1 recoded as Single/Married",
                   "16 states",
                   "Northern, Southern, Eastern, Western, Central",
                   "15 professions",
                   "18 product categories",
                   "1-4",
                   "188-23952")

# Generate table of variables
table_1 <- as.table(cbind(colnames(diwali_sales), diwali_class, diwali_desc,
                          diwali_values))
colnames(table_1) <- c("Variable", "Class", "Description", "Values")
row.names(table_1) <- seq(1:13)

# Generate a nice table with kable
my_table <- kable(table_1, "html") %>%
  kable_styling(full_width = FALSE)

# Display the table
my_table

# Summary checks
# summary(diwali_sales[which(diwali_sales$Age.Group == "0-17"), "Marital_Status"])
# summary(diwali_sales[which(diwali_sales$Age == 12), "Marital_Status"])
# unique(diwali_sales$State)
# unique(diwali_sales$Zone)
# unique(diwali_sales$Product_Category)
```

## What is the most prevalent customer group for each product category?

### Introduction
A natural direction when looking at transactional data is to find which
demographic engages in the highest number of transactions for any given
business. Finding these patterns can be helfpul for businesses in increasing
revenue in multiple ways such as knowing which products to focus on, which kind
of customers to target, or highlighting potentially untapped customer groups.
In order to answer this question, features `Gender`, `Marital_Status`,
`Product_Category`, `Orders`, and `Amount` are used. `Orders` and `Amount` are
specifically used to address one potential confounding factor by calculating
`Price`.

### Approach
A barplot was used to assess the relationship between product category and
its interest within a customer demographic. This plot visualizes the number of
transactions for each product category, differences in gender using different
bar colors, and marital status using facet wraps. The count of each transaction
is particularly important since any sale leads to profit regardless of order
size. The barplot's visualizations also help give us both an absolute and
relevant gauge of interest as the lengths of the bars can be compared to each
other.

It is reasonable to believe that product prices could affect the number of
transactions and be a confounding factor for customer demographic analysis.
To assess if product price is introducing potential effects on the number of
sales, a boxplot of product category prices is generated. The boxplot provides
several important statistics such as the interquartile range that gives a
clearer picture of how similar or different item prices between categories are.
Because prices are likely to remain the same regardless of the customer, gender
and marital status are not utilized for this plot.

### Analysis
```{r}
#| label: question1-data-wrangling
#| warning: false

# Read in project data
diwali <- read.csv("data/diwali_sales_data.csv")

# Remove NA values
diwali <- diwali[-which(is.na(diwali$Amount)),]

# Remove lowest age group due to inherent issues
diwali <- diwali[-which(diwali$Age.Group == "0-17"),]

# Recode marital status in case age value is illegal
diwali <- diwali |>
  mutate(
    Marital_Status = case_when(
      Age < 21 & Gender =="M" ~ 0,
      TRUE ~ Marital_Status
    )
  )

# Modify dataset according to project analysis
diwali <- diwali |>
  mutate(
    # Refactorizes the product category based on similarity between products
    Product_Category = case_when(
      Product_Category %in% c("Footwear & Shoes") ~ "Clothing",
      Product_Category %in% c("Clothing & Apparel") ~ "Clothing",
      Product_Category %in% c("Books") ~ "Entertainment",
      Product_Category %in% c("Electronics & Gadgets") ~ "Entertainment",
      Product_Category %in% c("Games & Toys") ~ "Entertainment",
      Product_Category %in% c("Sports Products") ~ "Entertainment",
      Product_Category %in% c("Decor") ~ "Home",
      Product_Category %in% c("Furniture") ~ "Home",
      Product_Category %in% c("Hand & Power Tools") ~ "Home",
      Product_Category %in% c("Household items") ~ "Home",
      Product_Category %in% c("Tupperware") ~ "Home",
      Product_Category %in% c("Stationery") ~ "Office",
      Product_Category %in% c("Veterinary") ~ "Pet Care",
      TRUE ~ Product_Category
      ),
    
    # Refactorizes the marital status to detail the meaning
    Marital_Status = case_when(
      Marital_Status %in% 1 ~ "Married",
      Marital_Status %in% 0 ~ "Single"
      ),
    
    # Calculate the price of item within each transaction
    Price = Amount/Orders
  )

# Sort Gender factor level
diwali$Gender <- factor(diwali$Gender, levels = c("M", "F"))

# Sort Marital_Status factor level
diwali$Marital_Status <- factor(diwali$Marital_Status,
                                levels = c("Single", "Married"))
```

```{r} 
#| label: question1-barplot
#| warning: false
#| fig-align: "center"

# Check the number of product categories for refactor level
category_order <- as.data.frame(matrix(nrow = 8, ncol = 2))
category_order[,1] <- c("Auto", "Beauty", "Clothing", "Entertainment", "Food",
                    "Home", "Office", "Pet Care")
category_order[,2] <- as.integer(summary(as.factor(diwali$Product_Category)))
category_vector <- (category_order[order(category_order[,2]),1])

# Refactorizes the product category to sort the groups by count
diwali$Product_Category <- as.factor(diwali$Product_Category)
diwali$Product_Category <- factor(diwali$Product_Category,
                                  levels = category_vector)

# Bar plot showing the number of customers for product category across gender
# and marital status
ggplot(data = diwali, aes(y = Product_Category, fill = Gender)) +
  geom_bar(position = "dodge", width = 0.7) +
  scale_fill_manual(values = c("dodgerblue", "deeppink1")) +
  # Reorder legend to match the order of the bar plot
  guides(fill = guide_legend(reverse = TRUE)) +
  ggtitle("Total Count of Sale Type",
          "Across gender and marital status") +
  facet_wrap(~Marital_Status, ncol = 1) +
  ylab("Product Category") +
  xlab("Counts")

```

```{r}
#| label: question1-boxplot
#| fig-align: "center"

# Check the average order price of product category
diwali_price <- diwali |>
  group_by(Product_Category) |>
  summarize("Price" = median(Amount/Orders))
diwali_price <- as.data.frame(diwali_price)
category_vector <- diwali_price[order(diwali_price[,2]),1]

# Refactorizes the product category to sort the groups by count
diwali$Product_Category <- factor(diwali$Product_Category,
                                  levels = rev(category_vector))

# Boxplot of average order values per category
ggplot(data = diwali,
       aes(x = Product_Category, y = Amount/Orders)) +
  geom_boxplot() +
  scale_y_continuous(labels = label_number(scale = 1e-3, suffix = "K")) +
  guides(fill = guide_legend(reverse = TRUE)) +
  ggtitle("Product Price Range",
          "Ordered by median order price") +
  xlab("Product Category") +
  ylab("Order Price (Rupee)")
```

### Discussion
The barplot indicated that the number of female customers was more than double
the number of male customers for most product categories. This difference could
be attributed to a difference in culture in which men are expected to purchase
corporate gifts (sweets and souvenirs) for business associates while women are
expected to purchase gifts for families and friends, who are more likely to
desire the specified product categories in this dataset. Additionally, customers
who were single made more transactions than customers who were married across
all product categories. This could be because single people only need to
consider their own purchases and not others in the family. The majority of
transactions were made with clothing, entertainment, and food. Considering the
transactions are for the festival, it is likely that local food, games, and
regional clothing would be highlights of the festival.

The boxplot shows that prices across product categories were within the whisker
range aside from auto and pet care. Median values for each product category did
not have a pattern consistent with the barplot. These two details suggest that
order price does not have a confounding effect on the transaction numbers of
our subgroups.

An interesting detail to note is that auto had the largest range and highest
median price, which is likely due to the range of potential items associated
with auto such as car accessories, auto parts, or services.

## How does age affect a customer's total spending during the festival?

### Introduction
When analyzing customer demographics, age is another feature that is often
focused on since age is often associated with several other features such as
generational differences, employment status, disposable income, etc. Because
of potential changes in spending habits, we are interested in seeing how age
can affect how much a customer will spend at the festival. In addition to
`Gender` and `Marital_Status` used in the previous question, we also look at
`Age`, `Age.Group`, `User_ID`, and `Amount`. `Age` and `Age.Group` are used as
the independent variables of our plots. `User_ID` and `Age` are used to group
customers so that `Amount` can be summed to determine the amount a customer
spends at that year's festival (`Total`).

### Approach
We plan on assessing how age is related to spending total when age is either
a categorical variable (age group) or continuous variable (age). Performing the
analysis using age group gives more insight to spending habits of customers
based on maturity. Gender and marital status are also assessed since results
from the previous question have shown an impact of these features on the number
of transactions and could have a possible impact on total spending.

A boxplot is implemented when age group is used. This visualization is useful to
show several important statistical values such as the interquartile
range. Comparison of the total spending ranges across customers at
different age groups is visualized by simply comparing the boxes horizontally.
In addition, a regression is implemented using age as a continuous variable.
This visualization is useful in directly mapping the relationship between a
customer's age and total spending.

### Analysis
```{r}
#| label: question2-data-wrangling

# Modify data for Q2
diwali.Q2 <- diwali |>
  group_by(User_ID, Age) |>
  summarize(
    "Gender" = first(Gender),
    "Age_Group" = first(Age.Group),
    "Marital_Status" = first(Marital_Status),
    "Total" = sum(Amount))
```

```{r} 
#| label: question2-boxplot
#| warning: false
#| fig-align: "center"

# Boxplot to show customer spending within age group
ggplot(data = diwali.Q2,
       aes(x = Age_Group, y = Total, fill = Gender)) +
  geom_boxplot() +
  scale_y_continuous(labels = label_number(scale = 1e-3, suffix = "K")) +
  scale_fill_manual(values = c("dodgerblue", "deeppink1")) +
  coord_cartesian(ylim = c(0, 80000)) +
  facet_wrap(~Marital_Status, ncol = 2) +
  guides(fill = guide_legend(reverse = TRUE)) +
  ggtitle("Spending Range of Customer Demographic", "By age group") +
  xlab("Age Group") +
  ylab("Customer Spending (Rupees)")
```

```{r}
#| label: question2-regression
#| fig-align: "center"
#| 
# Sort Gender factor level
diwali.Q2$Gender <- factor(diwali.Q2$Gender, levels = c("M", "F"))

# Plot Q2 data
ggplot(data = diwali.Q2, aes(x = Age, y = Total, color = Gender)) +
  geom_smooth(linewidth = 2)  +
  scale_y_continuous(labels = label_number(scale = 1e-3, suffix = "K")) +
  facet_wrap(~Marital_Status, ncol = 2) +
  scale_color_manual(values = c("dodgerblue", "deeppink1")) +
  guides(color = guide_legend(reverse = TRUE)) +
  ggtitle("Customer Demographic Spending Insights",
          "Over age") +
  ylab("Customer Spending (Rupees)")
```

```{r}
#| label: question2-slope
#| output: false
# Split data into two subsets
diwali_single <- diwali.Q2[which(diwali.Q2$Marital_Status == "Single" &
                               diwali.Q2$Gender == "M"),]

diwali_married <- diwali.Q2[which(diwali.Q2$Marital_Status == "Married" &
                               diwali.Q2$Gender == "M"),]

# Generate linear models
single_lm <- lm(Total~Age, data = diwali_single)
married_lm <- lm(Total~Age, data = diwali_married)

# Round values of tidy objects
tidy_single <- tidy(single_lm)
tidy_single[,2:ncol(tidy_single)] <- round(tidy_single[, 2:ncol(tidy_single)], 2)
tidy_married <- tidy(married_lm)
tidy_married[,2:ncol(tidy_married)] <- round(tidy_married[, 2:ncol(tidy_married)], 2)

# Generate a nice table with kable
single_table <- kable(tidy_single, "html") %>%
  kable_styling(full_width = FALSE)

# Generate a nice table with kable
married_table <- kable(tidy_married, "html") %>%
  kable_styling(full_width = FALSE)

# Display the table
single_table
married_table
```

Single Men

```{r}
#| label: question2-single
#| echo: false

# Display the table
single_table
```

Married Men

```{r}
#| label: question2-married
#| echo: false

# Display the table
married_table
```

### Discussion
The boxplot shows that interquartile range of the total rupees spent at the
festival does not fluctuate greatly across age groups regardless of gender and
marital status. The consistency of values could be the result of customers
having a perception of how much spending is normal that does not greatly deviate
across age. Female customers appear to have larger outlier values and higher
prevalence compared to their male counterparts across almost every subgroup.
We, as a group, refuse to put forward a reason for this in absence of any clear
or known substantiation for this behavior.

From the regression, it is noticeable male spending is consistent across all
ages regardless of marital status. The consistent spending could be attributed
to the fact that festivals are more driven by women. Spending of female
customers peaks approximately at age 50. The shift in spending could be
attributed to disposable income in relation to the age of dependent family members.

Due to the flat linear appearance of men's spending, slope analysis was
performed for both single and married men. The p-values were larger than
$\alpha$ = 0.05 for both. This indicates that men's spending is unaffected by age.

## Limitations
Limitations of the dataset included the following:

-   Date of transaction not included

-   Inconsistent occupation data within same year

-   No key to product ID

-   States is not pan-India

-   Difficult to assess confounding due to lack of data

## References

Diwali Sales Repository:

<a href="https://github.com/rfordatascience/tidytuesday/tree/master/data/2023/2023-11-14">https://github.com/rfordatascience/tidytuesday/tree/master/data/2023/2023-11-14</a>

