---
title: "Proposal title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
# Load data here
df <- read.csv(
  'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-11-14/diwali_sales_data.csv')

str(df)
```

Diwali stands as a paramount cultural festival celebrated fervently by diverse
communities worldwide, and many sales are involved as part of the festivites. As
a topic of interest, it is worth looking at the demographics regarding sales and
what products are popular. The Diwali dataset comes from TidyTuesday 2023 and
has 13 factors and 11251 observations. Variables include the following:

User_ID
Cust_name
Product_ID
Gender
Age Group
Age
MArital_Status
State
Zone
Occuptation
Product_Category
Orders
Amount

## Questions

Q1: For each product category, what is the largest customer base based on gender
and marital status.

Q2: How much does a customer spend given their gender and age group.

## Analysis plan

Q1: Transform the Product_Category based on commonality of certain categories
(ie pet service and veterinary). Marital status is recoded such that 0 is
married and 1 is single. Generate bar plot the product category with dodge
position based on gender. Add a facet wrap based on marital status.

Q2: Transform the data by grouping the observations by customer ID and sum the 
spending amount. Afterwards we will generate a boxplot using the amount as the
response variable. The x-axis will be the age group. The gender could be used
as either a facet wrap or use the dodge position. This will be determined later.
